---
title: "Quarto Emails (with Connect): Including plots"
author: "Lisa Anders"
date: today
format: email
---

## Introduction

Emails can be sent using Quarto that include plots. Learn how to make this magic happen: <https://docs.posit.co/connect/user/quarto/#email-customization> 

This example queries the Public Package Manager API endpoint in order to gather usage metrics and sends those to the named users added to the content after publishing on Connect. 

## Generate contents

Generate contents: Package Manager metrics 

> Please note that most of the above mentioned API Endpoints are not officially documented and not fully supported and hence can be subject to change in future versions of PPM. Any of those Endpoints however are visible via using Developer Tools in the web browser.

Reference the documentation on the server API either on your URL, or using the publicly available instance: <https://packagemanager.posit.co/__api__/swagger/index.html> 

For the hidden endpoints I'm referencing the proprietary source code: <https://github.com/rstudio/package-manager/tree/main/docs/api> and <https://github.com/rstudio/package-manager/tree/main/src/api> 

Alternatively to the below, we could use the package that Joe has made for us: 

```r
devtools::install_git("https://github.com/rstudio/ppmtools")
library(ppmtools)
configure("https://colorado.posit.co/rspm")
get_package_count_history("archer", start_date="2024-05-01")
```

```{r echo = FALSE, results = 'hide', message=FALSE, warning=FALSE}
library(httr)
library(tidyr)
library(dplyr)
library(reactable)
library(DT)
library(stringr)

var_url = "https://packagemanager.posit.co/"
var_sources_internal_packages = c(161,9,5) # source id's for transaction information c(8,9,15,161,177)
var_repos_internal_packages = 8 # source id's for transaction information c(8,9,15,161,177)
var_repos_bioconductor = 4 # repository id for bioconductor #13
var_version_bioconductor = 3.16
lookback = 30 #days
```

#### Server Status

```{r}
#Server status 
resp <- GET(paste0(var_url, "__api__/status"), format="json")
status <- content(resp, type="application/json")

rspm_version <- status$version
rspm_version
```

### Package Downloads 

PPM has an endpoint metrics/packages that allows you to get the download count of a given package for the last x days which should make it simple to expand this to date ranges. A sample call for package clustermq ’s download stats for the last 112 days would look like

`curl "https://packagemanager.posit.co/__api__/metrics/packages?_sort=count&_order=desc&_days=112&_sourceType=r&name_like=clustermq"`

Even better would be to get the download stats for all R packages on PPM via

`curl "https://colorado.posit.co/rspm/__api__/metrics/packages?_sourceType=r&_days=112&_sort=count&_order=desc"`

If you want to extract the metrics between x and y days prior to today, you run the API call above twice and then simply subtract the numbers.

Total  sourceType r package downloads: 

```{r}
resp <- GET(paste0(var_url,"__api__/metrics/packages?_sourceType=r&_days=",lookback,"&_sort=count&_order=desc"), format="json")
metricsPackagesR <- content(resp, type="application/json")
metricsPackagesR <- unnest_wider(tibble::tibble(metricsPackagesR = metricsPackagesR), metricsPackagesR) 

metricsPackagesR_sum <- metricsPackagesR %>% 
  summarize(total = sum(count)) %>% 
  mutate(language = "R")
```

Total  sourceType python package downloads:

```{r}
resp <- GET(paste0(var_url,"__api__/metrics/packages?_sourceType=python&_days=",lookback,"&_sort=count&_order=desc"), format="json")
metricsPackagesPython <- content(resp, type="application/json")
metricsPackagesPython <- unnest_wider(tibble::tibble(metricsPackagesPython = metricsPackagesPython), metricsPackagesPython)

metricsPackagesPython_sum <- metricsPackagesPython %>%
  summarize(total = sum(count)) %>% 
  mutate(language = "Python")
```

### Comparing R to Python total downloads: 

```{r}
library(ggrepel)

metricsPackages <- metricsPackagesPython_sum %>% 
  rbind(metricsPackagesR_sum)
  
metricsPackages_plot <- ggplot(metricsPackages, aes(x=total, y=language, label = total)) +
  geom_bar(stat = "identity", color="blue", fill="skyblue", alpha=0.7) +
  coord_flip() + 
  geom_text(aes(label = language), position = position_stack(vjust = 0.5)) +
  geom_text(aes(label= total), position = position_dodge(0.9), vjust = 0) + 
  ggtitle("Public Package Manager: Total downloads by language") + 
          #subtitle = "R Downloads") +  
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin = margin(t = 20,  # Top margin
                             r = 50,  # Right margin
                             b = 40,  # Bottom margin
                             l = 10)) # Left margin
  
metricsPackages_plot
```

### Cumulative downloads by day

By day cumulative sourceType r package downloads: 

```{r}
library(plotly)
library(ggrepel)
#library(ggpmisc)

# Raw package data, cumulative from date
metricsPackagesR_daily_raw <- data.frame()
for(i in 1:lookback){
  resp <- GET(paste0(var_url,"__api__/metrics/packages?_sourceType=r&_days=",i,"&_sort=count&_order=desc"), format="json")
  
  metricsPackagesR_daily_tmp <- content(resp, type="application/json")
  metricsPackagesR_daily_tmp <- unnest_wider(tibble::tibble(metricsPackagesR_daily_tmp = metricsPackagesR_daily_tmp), metricsPackagesR_daily_tmp) %>% 
    mutate(lookback = i, date = Sys.Date()-lookback) %>%
    mutate(language ="R")
  
metricsPackagesR_daily_raw <- rbind(metricsPackagesR_daily_raw, metricsPackagesR_daily_tmp)
}

# Reverse engineer dailies from reverse cumulative 
metricsPackagesR_daily_raw <- metricsPackagesR_daily_raw %>% 
  group_by(name) %>%
  mutate(daily_count = c(count[1],diff(count)))

# Summary
metricsPackagesR_daily <- metricsPackagesR_daily_raw %>% group_by(lookback, date, language) %>%
  summarize(downloads = sum(daily_count)) %>%
  group_by(language) %>%
  arrange(date) %>% 
  mutate(cumulative_sum = cumsum(downloads))

# Area chart
plot_r_package_downloads <- ggplot(metricsPackagesR_daily) + 
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks = "1 day") + 
  # Daily line
  geom_line(aes(x = date, y = downloads), 
            color = 4,      # Line color
            lwd = 0.5,      # Line width
            linetype = 1) + # Line type 
  geom_point(aes(x = date, y = downloads)) + 
  # Labels defined here
  geom_label_repel(
    data = metricsPackagesR_daily %>% filter(date == max(date)), 
    aes(date, downloads, label = "daily downloads"),
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.4, "lines"),
    color = "black",
    fill = "white") + 
  # Cumulative line 
  geom_line(aes(x = date, y = cumulative_sum), 
            color = 2,
            lwd = 1,  
            linetype = "dashed") +
  geom_point(aes(x = date, y = cumulative_sum)) + 
  ggtitle("Public Package Manager: R Downloads") + 
          #subtitle = "R Downloads") +  
  # Labels defined here
  geom_label_repel(
    data = metricsPackagesR_daily %>% filter(date == max(date)), 
    aes(date, cumulative_sum, label = "cumulative downloads"),
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.4, "lines"),
    color = "black",
    fill = "white") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin = margin(t = 20,  # Top margin
                             r = 50,  # Right margin
                             b = 40,  # Bottom margin
                             l = 10)) # Left margin
  
plot_r_package_downloads
```

By day cumulative sourceType python package downloads:

```{r}
# Raw package data, cumulative from date
metricsPackagesPython_daily_raw <- data.frame()
for(i in 1:lookback){
  resp <- GET(paste0(var_url,"__api__/metrics/packages?_sourceType=python&_days=",i,"&_sort=count&_order=desc"), format="json")
  
  metricsPackagesPython_daily_tmp <- content(resp, type="application/json")
  metricsPackagesPython_daily_tmp <- unnest_wider(tibble::tibble(metricsPackagesPython_daily_tmp = metricsPackagesPython_daily_tmp), metricsPackagesPython_daily_tmp) %>% 
    mutate(lookback = i, date = Sys.Date()-lookback) %>%
    mutate(language ="Python")
  
metricsPackagesPython_daily_raw <- rbind(metricsPackagesPython_daily_raw, metricsPackagesPython_daily_tmp)
}

# Reverse engineer dailies from reverse cumulative 
metricsPackagesPython_daily_raw <- metricsPackagesPython_daily_raw %>% 
  group_by(name) %>%
  mutate(daily_count = c(count[1],diff(count)))

# Summary
metricsPackagesPython_daily <- metricsPackagesPython_daily_raw %>% group_by(lookback, date, language) %>%
  summarize(downloads = sum(daily_count)) %>%
  group_by(language) %>%
  arrange(date) %>% 
  mutate(cumulative_sum = cumsum(downloads))

# Area chart
plot_Python_package_downloads <- ggplot(metricsPackagesPython_daily) + 
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks = "1 day") + 
  # Daily line
  geom_line(aes(x = date, y = downloads), 
            color = 4,      # Line color
            lwd = 0.5,      # Line width
            linetype = 1) + # Line type 
  geom_point(aes(x = date, y = downloads)) + 
  # Labels defined here
  geom_label_repel(
    data = metricsPackagesPython_daily %>% filter(date == max(date)), 
    aes(date, downloads, label = "daily downloads"),
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.4, "lines"),
    color = "black",
    fill = "white") + 
  # Cumulative line 
  geom_line(aes(x = date, y = cumulative_sum), 
            color = 2,
            lwd = 1,  
            linetype = "dashed") +
  geom_point(aes(x = date, y = cumulative_sum)) + 
  # Labels defined here
  geom_label_repel(
    data = metricsPackagesPython_daily %>% filter(date == max(date)), 
    aes(date, cumulative_sum, label = "cumulative downloads"),
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.4, "lines"),
    color = "black",
    fill = "white") + 
  ggtitle("Public Package Manager: Python Downloads") + 
          #subtitle = "Python Downloads") +  
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin = margin(t = 20,  # Top margin
                             r = 50,  # Right margin
                             b = 40,  # Bottom margin
                             l = 10)) # Left margin

#plot_Python_package_downloads_plotly <- ggplotly(plot_Python_package_downloads) 
plot_Python_package_downloads
```

### Stack of individual package downloads, top 10

R individual package downloads daily stacked: 

```{r}
library(paletteer)
library(patchwork)
library(gridExtra)

# Only top 10 show up in label 
metricsPackagesR_daily_raw_top10 <- metricsPackagesR_daily_raw %>% 
  group_by(language, name) %>%
  summarize(downloads = sum(daily_count)) %>% 
  arrange(desc(downloads)) %>%
  ungroup() %>%
  mutate(name_label = ifelse(row_number() <= 10, name, "other")) %>%
  arrange(downloads) %>%
  mutate(name_label = factor(name_label, levels=unique(name_label))) 
  
metricsPackagesR_daily_raw_plot_data <- metricsPackagesR_daily_raw %>%
  merge(metricsPackagesR_daily_raw_top10 %>% select(name, name_label, downloads), by = "name") %>% 
  arrange(downloads)

# Plot daily stack
metricsPackagesR_daily_raw_plot <- ggplot(metricsPackagesR_daily_raw_plot_data, aes(fill=name_label, y=daily_count, x=date)) + 
  geom_bar(position='stack', stat='identity') + 
  ggtitle("Public Package Manager R Packages: ") +
          #subtitle = "R Downloads") +
  scale_color_paletteer_d("colorBlindness::paletteMartin") + 
  scale_colour_paletteer_d("colorBlindness::paletteMartin") + 
  scale_fill_paletteer_d("colorBlindness::paletteMartin") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        plot.margin = margin(t = 20,  # Top margin
                             r = 50,  # Right margin
                             b = 40,  # Bottom margin
                             l = 10)) # Left margin

# Plot top 10 
metricsPackagesR_daily_raw_plot2 <- ggplot(metricsPackagesR_daily_raw_plot_data %>% select(name, name_label, downloads) %>% unique(), aes(fill=name_label, y=downloads, x=name_label)) + 
  geom_bar(position='stack', stat='identity') + 
  ggtitle("Downloads") +
          #subtitle = "R Downloads") +
  scale_color_paletteer_d("colorBlindness::paletteMartin") + 
  scale_colour_paletteer_d("colorBlindness::paletteMartin") + 
  scale_fill_paletteer_d("colorBlindness::paletteMartin") + 
  coord_flip(clip = "off") + 
  # geom_text(data = metricsPackagesR_daily_raw_plot_data %>% select(name, name_label, downloads) %>% unique() %>% top_n(5), aes(x = name_label, y = downloads, label = downloads), position = position_stack(vjust = 0.5)) +
  geom_text(data = metricsPackagesR_daily_raw_plot_data %>% select(name, name_label, downloads) %>% unique() %>% top_n(5), aes(x = name_label, y = downloads, label = downloads), position = position_dodge(width=0.9), hjust = -0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        panel.border = element_rect(linetype = "dashed", fill = NA, colour = "white"),
        axis.title.y = element_blank(),
        plot.margin = margin(t = 20,  # Top margin
                             r = 50,  # Right margin
                             b = 40,  # Bottom margin
                             l = 10)) # Left margin

metricsPackagesR_daily_raw_plot
metricsPackagesR_daily_raw_plot2

metricsPackagesR_daily_raw_plot3 <- gridExtra::grid.arrange(metricsPackagesR_daily_raw_plot,metricsPackagesR_daily_raw_plot2, ncol=2, widths = 3:2)

metricsPackagesR_daily_raw_plot3
```

Python individual package downloads daily stacked: 

```{r}
library(paletteer)
library(patchwork)
library(gridExtra)

# Only top 10 show up in label 
metricsPackagesPython_daily_raw_top10 <- metricsPackagesPython_daily_raw %>% 
  group_by(language, name) %>%
  summarize(downloads = sum(daily_count)) %>% 
  arrange(desc(downloads)) %>%
  ungroup() %>%
  mutate(name_label = ifelse(row_number() <= 10, name, "other")) %>%
  arrange(downloads) %>%
  mutate(name_label = factor(name_label, levels=unique(name_label))) 
  
metricsPackagesPython_daily_raw_plot_data <- metricsPackagesPython_daily_raw %>%
  merge(metricsPackagesPython_daily_raw_top10 %>% select(name, name_label, downloads), by = "name") %>% 
  arrange(downloads)

# Plot daily stack
metricsPackagesPython_daily_raw_plot <- ggplot(metricsPackagesPython_daily_raw_plot_data, aes(fill=name_label, y=daily_count, x=date)) + 
  geom_bar(position='stack', stat='identity') + 
  ggtitle("Public Package Manager Python Packages: ") +
          #subtitle = "R Downloads") +
  scale_color_paletteer_d("colorBlindness::paletteMartin") + 
  scale_colour_paletteer_d("colorBlindness::paletteMartin") + 
  scale_fill_paletteer_d("colorBlindness::paletteMartin") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        plot.margin = margin(t = 20,  # Top margin
                             r = 50,  # Right margin
                             b = 40,  # Bottom margin
                             l = 10)) # Left margin

# Plot top 10 
metricsPackagesPython_daily_raw_plot2 <- ggplot(metricsPackagesPython_daily_raw_plot_data %>% select(name, name_label, downloads) %>% unique(), aes(fill=name_label, y=downloads, x=name_label)) + 
  geom_bar(position='stack', stat='identity') + 
  ggtitle("Downloads") +
          #subtitle = "R Downloads") +
  scale_color_paletteer_d("colorBlindness::paletteMartin") + 
  scale_colour_paletteer_d("colorBlindness::paletteMartin") + 
  scale_fill_paletteer_d("colorBlindness::paletteMartin") + 
  coord_flip(clip = "off") + 
  # geom_text(data = metricsPackagesPython_daily_raw_plot_data %>% select(name, name_label, downloads) %>% unique() %>% top_n(5), aes(label = downloads), position = position_stack(vjust = 0.5)) +
  geom_text(data = metricsPackagesPython_daily_raw_plot_data %>% select(name, name_label, downloads) %>% unique() %>% top_n(5), aes(x = name_label, y = downloads, label = downloads), position = position_dodge(width=0.9), hjust = -0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.border = element_rect(linetype = "dashed", fill = NA, colour = "white"),
        legend.position = "none",
        axis.title.y = element_blank(),
        plot.margin = margin(t = 20,  # Top margin
                             r = 50,  # Right margin
                             b = 40,  # Bottom margin
                             l = 10)) # Left margin

metricsPackagesPython_daily_raw_plot
metricsPackagesPython_daily_raw_plot2

metricsPackagesPython_daily_raw_plot3 <- gridExtra::grid.arrange(metricsPackagesPython_daily_raw_plot,metricsPackagesPython_daily_raw_plot2, ncol=2, widths = 3:2)

metricsPackagesPython_daily_raw_plot3
```

## Send the generated metrics in an email

::: {.email}

::: {.subject}
Curated report of p3m metrics
:::

Lookback: `{r} lookback`

Package Manager is on version: `{r} rspm_version`

Package Manager total downloads: 

```{r}
#| echo: false

metricsPackages_plot
```

R Package downloads from p3m: 

```{r}
#| echo: false

plot_r_package_downloads
```

R Package downloads from p3m, daily stack and pareto: 

```{r}
#| echo: false

gridExtra::grid.arrange(metricsPackagesR_daily_raw_plot,metricsPackagesR_daily_raw_plot2, ncol=2, widths = 3:2)
```

Python Package downloads from p3m: 

```{r}
#| echo: false

plot_Python_package_downloads
```

Python Package downloads from p3m, daily stack and pareto: 

```{r}
#| echo: false

gridExtra::grid.arrange(metricsPackagesPython_daily_raw_plot,metricsPackagesPython_daily_raw_plot2, ncol=2, widths = 3:2)
```

This is what happens if you try to include an interactive plot: 

```{r}
#| echo: false

ggplotly(metricsPackages_plot)
```

:::