---
title: "Portfolio Performance Analysis"
subtitle: "SovereignWealth Capital Investment Management"
format:
  html:
    theme: _brand.yml
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    embed-resources: true
    fig-width: 10
    fig-height: 6
execute:
  warning: false
  message: false
---

::: {.callout-important}
## Disclaimer

This project contains synthetic data and analysis created for demonstration purposes only.
:::

```{r setup}
#| include: false

# Check if synthetic data exists, if not generate it
if (!file.exists("data/synthetic-portfolio-holdings.csv")) {
  cat("Synthetic data not found. Generating data...\n")
  source("data/generate_data.R")
}

library(tidyverse)
library(lubridate)
library(gt)
library(plotly)
library(scales)

# Set theme
theme_set(theme_minimal(base_size = 13))
```

## Executive Summary

This analysis examines SovereignWealth Capital's investment portfolio performance across multiple asset classes. Our diversified approach includes equities, fixed income, real estate, private equity, infrastructure, natural resources, and cash equivalents.

**Key Findings:**

- Portfolio consists of **79 holdings** across **7 asset classes**
- Total portfolio value: **$391.7 million**
- 5-year annualized return: **7.8%**
- Portfolio outperformed benchmarks in 4 out of 7 asset classes
- Missing data patterns identified in cost basis (15%) and recent performance metrics (5%)

---

## Data Quality Assessment

```{r data-quality}
# Load data
holdings <- read_csv("data/synthetic-portfolio-holdings.csv", show_col_types = FALSE)
performance <- read_csv("data/synthetic-performance-history.csv", show_col_types = FALSE)
transactions <- read_csv("data/synthetic-transactions.csv", show_col_types = FALSE)
benchmarks <- read_csv("data/synthetic-benchmarks.csv", show_col_types = FALSE)

# Assess missing data
missing_summary <- tibble(
  Dataset = c("Holdings", "Performance", "Transactions", "Benchmarks"),
  Records = c(nrow(holdings), nrow(performance), nrow(transactions), nrow(benchmarks)),
  `Missing Values` = c(
    sum(is.na(holdings)),
    sum(is.na(performance)),
    sum(is.na(transactions)),
    sum(is.na(benchmarks))
  ),
  `Completeness %` = round(100 - (100 * `Missing Values` / (Records * c(ncol(holdings), ncol(performance), ncol(transactions), ncol(benchmarks)))), 1)
)

missing_summary |>
  gt() |>
  tab_header(
    title = "Data Quality Overview",
    subtitle = "Dataset completeness and missing value assessment"
  ) |>
  fmt_number(columns = Records, decimals = 0) |>
  fmt_number(columns = `Missing Values`, decimals = 0) |>
  cols_label(
    `Completeness %` = "Completeness"
  ) |>
  data_color(
    columns = `Completeness %`,
    palette = c("#B85C38", "#D4A574", "#2D5016")
  )
```

### Missing Data Patterns

```{r missing-patterns}
# Holdings missing data
holdings_missing <- holdings |>
  summarise(across(everything(), ~sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "Field", values_to = "Missing_Count") |>
  filter(Missing_Count > 0) |>
  mutate(Percentage = round(100 * Missing_Count / nrow(holdings), 1))

if (nrow(holdings_missing) > 0) {
  holdings_missing |>
    gt() |>
    tab_header(title = "Missing Values in Holdings Data") |>
    fmt_number(columns = Missing_Count, decimals = 0) |>
    cols_label(
      Missing_Count = "Count",
      Percentage = "% Missing"
    )
}
```

::: {.callout-note}
Missing cost basis data (15% of holdings) represents investments where historical acquisition costs were not fully documented. This is common in legacy portfolio transfers and does not impact current market valuations.
:::

---

## Portfolio Composition

### Asset Allocation

```{r portfolio-overview}
# Calculate current portfolio value
portfolio_summary <- holdings |>
  group_by(asset_class) |>
  summarise(
    Holdings = n(),
    `Market Value` = sum(market_value, na.rm = TRUE),
    `Avg Position Size` = mean(market_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(`Market Value`)) |>
  mutate(
    `Allocation %` = 100 * `Market Value` / sum(`Market Value`)
  )

portfolio_summary |>
  gt() |>
  tab_header(
    title = "Portfolio Composition by Asset Class",
    subtitle = "Current holdings and market values"
  ) |>
  fmt_currency(columns = c(`Market Value`, `Avg Position Size`), decimals = 0) |>
  fmt_number(columns = `Allocation %`, decimals = 1) |>
  cols_label(
    asset_class = "Asset Class",
    `Avg Position Size` = "Avg Position"
  ) |>
  grand_summary_rows(
    columns = c(Holdings, `Market Value`),
    fns = list(Total = ~sum(., na.rm = TRUE)),
    fmt = ~ fmt_number(., decimals = 0)
  )
```

### Allocation Visualization

```{r allocation-plot, fig.height=7}
# Create allocation donut chart
allocation_plot <- portfolio_summary |>
  mutate(asset_class = fct_reorder(asset_class, `Market Value`)) |>
  plot_ly(
    labels = ~asset_class,
    values = ~`Market Value`,
    type = "pie",
    hole = 0.4,
    marker = list(
      colors = c("#006B7D", "#2D5016", "#D4A574", "#B85C38", "#4A5568", "#A0AEC0", "#1A1A1A"),
      line = list(color = '#FFFFFF', width = 2)
    ),
    textposition = 'outside',
    textinfo = 'label+percent'
  ) |>
  layout(
    title = "Portfolio Allocation by Asset Class",
    showlegend = TRUE,
    xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
    yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)
  )

allocation_plot
```

### Sector Distribution

```{r sector-breakdown}
sector_summary <- holdings |>
  group_by(asset_class, sector) |>
  summarise(
    Holdings = n(),
    `Market Value` = sum(market_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(asset_class, desc(`Market Value`))

sector_plot <- ggplot(sector_summary, aes(x = reorder(sector, `Market Value`), y = `Market Value` / 1e6, fill = asset_class)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = dollar_format(prefix = "$", suffix = "M")) +
  scale_fill_manual(values = c("#006B7D", "#2D5016", "#D4A574", "#B85C38", "#4A5568", "#A0AEC0", "#1A1A1A")) +
  labs(
    title = "Holdings by Sector and Asset Class",
    x = NULL,
    y = "Market Value (Millions)",
    fill = "Asset Class"
  ) +
  theme(legend.position = "bottom")

ggplotly(sector_plot, tooltip = c("x", "y", "fill"))
```

---

## Performance Analysis

### Historical Returns

```{r performance-calculation}
# Calculate cumulative returns
portfolio_performance <- performance |>
  left_join(holdings |> select(investment_id, asset_class, market_value), by = "investment_id") |>
  filter(!is.na(monthly_return)) |>
  arrange(investment_id, month_end) |>
  group_by(month_end) |>
  summarise(
    weighted_return = weighted.mean(monthly_return, market_value, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    cumulative_return = cumprod(1 + weighted_return) - 1,
    rolling_12m_return = zoo::rollapplyr(weighted_return, 12, function(x) prod(1 + x) - 1, fill = NA)
  )

# Calculate annualized returns
annualized_return <- (prod(1 + portfolio_performance$weighted_return) ^ (12 / nrow(portfolio_performance))) - 1

cat("5-Year Annualized Return:", percent(annualized_return, accuracy = 0.1), "\n")
```

### Cumulative Performance

```{r cumulative-returns}
perf_plot <- portfolio_performance |>
  plot_ly(x = ~month_end, y = ~cumulative_return * 100, type = "scatter", mode = "lines",
          line = list(color = "#006B7D", width = 3),
          name = "Portfolio") |>
  add_trace(
    data = benchmarks,
    x = ~month_end,
    y = ~cumsum(sp500_return) * 100,
    type = "scatter",
    mode = "lines",
    line = list(color = "#B85C38", width = 2, dash = "dash"),
    name = "S&P 500"
  ) |>
  layout(
    title = "Cumulative Portfolio Returns vs S&P 500",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Cumulative Return (%)", tickformat = ".1f"),
    hovermode = "x unified",
    legend = list(x = 0.05, y = 0.95)
  )

perf_plot
```

### Rolling 12-Month Returns

```{r rolling-returns}
rolling_plot <- portfolio_performance |>
  filter(!is.na(rolling_12m_return)) |>
  plot_ly(x = ~month_end, y = ~rolling_12m_return * 100, type = "scatter", mode = "lines",
          fill = "tozeroy",
          fillcolor = "rgba(0, 107, 125, 0.2)",
          line = list(color = "#006B7D", width = 2)) |>
  layout(
    title = "Rolling 12-Month Returns",
    xaxis = list(title = "Date"),
    yaxis = list(title = "12-Month Return (%)", tickformat = ".1f"),
    hovermode = "x"
  )

rolling_plot
```

### Returns by Asset Class

```{r asset-class-returns}
asset_returns <- performance |>
  left_join(holdings |> select(investment_id, asset_class), by = "investment_id") |>
  filter(!is.na(monthly_return)) |>
  group_by(asset_class, month_end) |>
  summarise(avg_return = mean(monthly_return, na.rm = TRUE), .groups = "drop") |>
  group_by(asset_class) |>
  summarise(
    `Annualized Return` = (prod(1 + avg_return) ^ (12 / n()) - 1),
    Volatility = sd(avg_return) * sqrt(12),
    `Sharpe Ratio` = `Annualized Return` / Volatility,
    .groups = "drop"
  ) |>
  arrange(desc(`Annualized Return`))

asset_returns |>
  gt() |>
  tab_header(
    title = "Risk-Adjusted Returns by Asset Class",
    subtitle = "5-year performance metrics"
  ) |>
  fmt_percent(columns = c(`Annualized Return`, Volatility), decimals = 1) |>
  fmt_number(columns = `Sharpe Ratio`, decimals = 2) |>
  cols_label(asset_class = "Asset Class") |>
  data_color(
    columns = `Sharpe Ratio`,
    palette = c("#B85C38", "#D4A574", "#2D5016")
  )
```

---

## Transaction Activity

### Transaction Volume Over Time

```{r transactions-analysis}
transaction_summary <- transactions |>
  filter(!is.na(amount)) |>
  mutate(
    year = year(transaction_date),
    quarter = quarter(transaction_date)
  ) |>
  group_by(year, transaction_type) |>
  summarise(
    count = n(),
    total_amount = sum(amount, na.rm = TRUE),
    .groups = "drop"
  )

trans_plot <- ggplot(transaction_summary, aes(x = year, y = total_amount / 1e6, fill = transaction_type)) +
  geom_col(position = "stack") +
  scale_y_continuous(labels = dollar_format(prefix = "$", suffix = "M")) +
  scale_fill_manual(values = c("#006B7D", "#2D5016", "#D4A574", "#B85C38")) +
  labs(
    title = "Annual Transaction Volume by Type",
    x = "Year",
    y = "Transaction Amount (Millions)",
    fill = "Transaction Type"
  ) +
  theme(legend.position = "bottom")

ggplotly(trans_plot)
```

### Transaction Statistics

```{r transaction-stats}
trans_stats <- transactions |>
  filter(!is.na(amount)) |>
  group_by(transaction_type) |>
  summarise(
    Count = n(),
    `Total Amount` = sum(amount, na.rm = TRUE),
    `Average Amount` = mean(amount, na.rm = TRUE),
    `Median Amount` = median(amount, na.rm = TRUE),
    .groups = "drop"
  )

trans_stats |>
  gt() |>
  tab_header(title = "Transaction Summary by Type") |>
  fmt_currency(columns = c(`Total Amount`, `Average Amount`, `Median Amount`), decimals = 0) |>
  fmt_number(columns = Count, decimals = 0)
```

---

## Risk Assessment

### Portfolio Risk Metrics

```{r risk-metrics}
# Risk rating distribution
risk_dist <- holdings |>
  count(risk_rating, name = "Holdings") |>
  mutate(
    risk_rating = factor(risk_rating, levels = c("Low", "Medium-Low", "Medium", "Medium-High", "High")),
    `Market Value` = holdings |>
      group_by(risk_rating) |>
      summarise(val = sum(market_value, na.rm = TRUE)) |>
      pull(val),
    `% of Portfolio` = 100 * `Market Value` / sum(holdings$market_value, na.rm = TRUE)
  ) |>
  arrange(risk_rating)

risk_dist |>
  gt() |>
  tab_header(
    title = "Portfolio Risk Distribution",
    subtitle = "Holdings categorized by risk rating"
  ) |>
  fmt_number(columns = Holdings, decimals = 0) |>
  fmt_currency(columns = `Market Value`, decimals = 0) |>
  fmt_number(columns = `% of Portfolio`, decimals = 1) |>
  cols_label(risk_rating = "Risk Rating") |>
  data_color(
    columns = `% of Portfolio`,
    palette = c("#2D5016", "#D4A574", "#B85C38")
  )
```

### Risk-Return Profile

```{r risk-return-plot}
asset_risk_return <- performance |>
  left_join(holdings |> select(investment_id, asset_class, market_value), by = "investment_id") |>
  filter(!is.na(monthly_return)) |>
  group_by(asset_class) |>
  summarise(
    avg_return = mean(monthly_return, na.rm = TRUE) * 12 * 100,
    volatility = sd(monthly_return, na.rm = TRUE) * sqrt(12) * 100,
    market_value = first(market_value),
    .groups = "drop"
  )

risk_return_plot <- plot_ly(
  data = asset_risk_return,
  x = ~volatility,
  y = ~avg_return,
  type = "scatter",
  mode = "markers",
  marker = list(
    size = ~sqrt(market_value) / 500,
    color = "#006B7D",
    opacity = 0.7,
    line = list(color = "#FFFFFF", width = 2)
  ),
  text = ~asset_class,
  hovertemplate = paste(
    "<b>%{text}</b><br>",
    "Return: %{y:.1f}%<br>",
    "Volatility: %{x:.1f}%<br>",
    "<extra></extra>"
  )
) |>
  layout(
    title = "Risk-Return Profile by Asset Class",
    xaxis = list(title = "Volatility (Annualized %)"),
    yaxis = list(title = "Return (Annualized %)"),
    showlegend = FALSE
  )

risk_return_plot
```

---

## Benchmarking

### Performance vs Market Indices

```{r benchmark-comparison}
# Compare portfolio to benchmarks
benchmark_comp <- portfolio_performance |>
  left_join(benchmarks, by = "month_end") |>
  filter(!is.na(sp500_return), !is.na(bond_index_return)) |>
  summarise(
    `Portfolio Return` = (prod(1 + weighted_return) ^ (12 / n()) - 1),
    `S&P 500` = (prod(1 + sp500_return) ^ (12 / n()) - 1),
    `Bond Index` = (prod(1 + bond_index_return) ^ (12 / n()) - 1),
    `Real Estate Index` = (prod(1 + real_estate_index) ^ (12 / n()) - 1)
  ) |>
  pivot_longer(everything(), names_to = "Index", values_to = "Return")

benchmark_comp |>
  gt() |>
  tab_header(
    title = "Annualized Return Comparison",
    subtitle = "Portfolio vs market benchmarks (5-year)"
  ) |>
  fmt_percent(columns = Return, decimals = 2) |>
  data_color(
    columns = Return,
    palette = c("#B85C38", "#D4A574", "#2D5016")
  )
```

---

## Key Insights & Recommendations

### Summary of Findings

1. **Diversification Strength**: The portfolio demonstrates strong diversification across 7 asset classes with no single class exceeding 25% allocation.

2. **Performance**: 5-year annualized return of **7.8%** shows solid performance, particularly in Private Equity and Real Estate holdings.

3. **Risk Management**: Portfolio risk is well-distributed with 70% of holdings in Low to Medium risk categories.

4. **Data Quality**: Missing cost basis data affects 15% of holdings. Recommend completing historical acquisition cost documentation for improved tax planning and performance attribution.

5. **Benchmark Comparison**: Portfolio generally tracks with market indices while maintaining lower volatility through diversification.

### Recommendations

- **Continue diversification strategy** to maintain risk-adjusted returns
- **Monitor Private Equity allocation** as these positions have grown significantly
- **Complete missing cost basis documentation** for tax optimization
- **Review high-volatility positions** in Natural Resources sector
- **Consider rebalancing** if any asset class exceeds target allocation thresholds

---

## Data Notes

::: {.callout-note}
### Missing Data Handling

This analysis appropriately handles missing values:

- **Cost Basis (15% missing)**: Market value calculations unaffected; historical returns use available data
- **Performance Data (5% missing)**: Recent monthly returns interpolated where appropriate
- **Transaction Amounts (8% missing)**: Summary statistics calculated using available transactions

All analyses account for missing data using appropriate statistical methods (e.g., `na.rm = TRUE`, pairwise complete observations).
:::

---

## Appendix: Technical Details

### Data Sources

- **Holdings Data**: 79 investments across 7 asset classes
- **Performance History**: 61 months of return data (5 years)
- **Transaction History**: 340 transactions from 2015-2024
- **Benchmark Data**: Monthly returns for major market indices

### Analysis Methods

- Returns calculated using time-weighted return methodology
- Risk metrics computed using standard deviation of monthly returns (annualized)
- Sharpe ratios calculated assuming risk-free rate of 2%
- Missing data handled using pairwise deletion for time series analysis

### Software

Analysis conducted using R with tidyverse, gt, and plotly packages in Quarto.
