---
title: "Email Document"
output: blastula::blastula_email
editor_options: 
  chunk_output_type: console
---

Breathe easy! AQI measurements are all good for `r nrow(good_air)` measurements.

![[^1]](blue_skies.jpeg){width="225"}

[^1]: Photo by <a href="https://unsplash.com/@frostroomhead">Rodion Kutsaiev</a> on <a href="https://unsplash.com/images/nature/cloud">Unsplash</a>
  

comprehensive view of local air quality, see the [Air Quality Dashboard](https://pub.current.posit.team/public/aqi_dashboard/).

A sample of the most recent readings from 6 random sites:

```{r echo=FALSE}
most_recent <- good_air %>% 
  group_by(station) %>% 
  arrange(desc(time)) |> 
  slice(1) %>% ungroup() %>% 
  select(station, us_aqi, time, color)

sample_readings <- most_recent %>% sample_n(6)

ggplot(sample_readings, aes(fill = color, ymax = us_aqi/350, ymin = 0, xmax = 2, xmin = 1)) +
  geom_rect(aes(ymax=1, ymin=0, xmax=2, xmin=1), fill ="whitesmoke") +
  geom_rect() + 
  coord_polar(theta = "y",start=-pi/2) + xlim(c(0, 2)) + ylim(c(0,2)) +
  geom_text(aes(x = 0, y = 0, label = us_aqi, colour=color), size=5.5) +
  geom_text(aes(x=1.5, y=1.5, label=station), size=3) + 
  facet_wrap(~station, ncol = 3) +
  theme_void() +
  scale_fill_manual(values = c(
    "green"="forestgreen","yellow"="yellow", "orange"="darkorange", 
    "red"="red", "purple"="purple1", "maroon"="red4")) +
  scale_colour_manual(values = c("green"="forestgreen","yellow"="yellow", "orange"="darkorange", 
                                 "red"="red", "purple"="purple1", "maroon"="red4")) +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank()) +
  guides(fill="none") +
  guides(colour="none")
```

```{r echo=FALSE}
suppressMessages(library(gt))
```

The highest reading over the last `r days_back` days is from the **`r max_site_data$station[1]`** site. This peak reading was **`r max_site_data$max_aqi[1]`**.

Here's the trend at that site:

```{r echo=FALSE}
site_plot
```

Here's a summary of the highest readings over the last `r days_back` days from each site:

```{r echo=FALSE}
good_air %>% select(station, us_aqi, us_aqi_pm2_5, us_aqi_pm10) %>% 
  group_by(station) %>% 
  arrange(desc(us_aqi)) %>% 
  slice(1) %>% 
  ungroup() %>%
  gt() 
```
